---
title: "Dartmouth_docsPCSA"
output:
  html_notebook: default
  github_document: default
date: '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(stringr); library(foreign); library(QuantPsyc); library(psych); library(knitr); library(tidyverse)
```

### Load zcta2000 -> zcta2010 crosswalk for NYC

```{r}
# NYC tract2010 -> zcta2010 xwalk
nyc_zcta_xwalk <- read_csv("../dropbox/capstone/crosswalks/tract2010_zcta2010_xwalk.csv", col_types = "ccd")
```

### Data downloaded from: https://datawarehouse.hrsa.gov/data/dataDownload/pcsa2010download.aspx

# 2000 Data


### Import all files, filter to nyc, clean names, and merge together

```{r}
# NYC tract2010 -> zcta2010 xwalk
nyc_zcta_xwalk <- read_csv("../dropbox/capstone/crosswalks/tract2010_zcta2010_xwalk.csv", col_types = "ccd")

import_filter_clean <- function(filename) {
  drop_vars <- c("state", "county", "tract", "areaname", "stusab", 
                 "intptlat", "intptlon", "landsqmi", "pcsa", "pcsa_l")
  
  read.dbf(str_interp("../dropbox/capstone/data_raw/${filename}.dbf"), as.is = TRUE) %>% 
    janitor::clean_names() %>% 
    mutate(tract10 = str_c(county, tract)) %>% 
    select(-one_of(drop_vars)) %>% 
    semi_join(nyc_zcta_xwalk, by = "tract10") %>% 
    as_tibble()
}

dart10_all <- c("t_103113_1", "t_103113_2", "t_cnm_np_122013") %>% 
  map(import_filter_clean) %>% 
  reduce(left_join, by = "tract10")

```

### Create all basic variables

```{r}
dart10_all2 <- dart10_all %>% 
  transmute(
    tract10 = tract10,
    
    fem15_44 = tp12i30 + tp12i31 + tp12i32 + tp12i33 + tp12i34 + tp12i36 + tp12i37 + tp12i38,
    
    # Nurse-practitioner data
    cnm_fte = t_cnmfte, 
    np_fte = t_npfte,
    
    # These 2 variables are not available in the tract file (only PCSA)
    # we need to figure out if there are other variables in the tract file 
    # that these are included under

    # famprac = tf_doc,
    # internist = ti_doc,
    pcphys = tg_doc,
    specialist = ts_doc,
    obgyn = to_doc,
    nonfedpcp = tg_nfeddoc,
    nonfedspec = ts_nfeddoc,
    nonfedob = to_nfeddoc,
    img_pcp = tg_img,
    img_spec = ts_img,
    img_ob = to_img,
    pa_pcp = tpa_pc,
    pa_spec = tpa_pc_3sp, # GP/IM/Pediatrics PAs - as specialists
    pa_ob = tpa_pc_ob, # ob pa
    pa_oth = tpa_pc_oth, # geriatrics or unk specialty PA
    partb_pcp = tvt_pc10, # pc visits Part b only
    partb_amb = tvt_am10, # ambulatory visits part b only
    edvt = tvt_ed10, # est total number of ED visits per day
    pcpvt_rhc = tvt_rhc10,
    pcpvt_fqhc = tvt_fqhc10,
    medicare_acscd = tacs_10, # acs discharges
    medicare_denom = tbene_n10, # total bene pop for 2010, all medicare, not part-specific
    pcpdenom = tvden_10, # denom for PCP rate
    totpop = tage00_14 + tage15_64 + tage65_up, 
    hospdenom = thden_10
    # The following variables are not available for the tract (only pcsa)
    # But I don't see them used anywhere so this is probably fine

    # ppop_pcp = tpop_gdoc,
    # acs_crudert = tcracs_10, # their calculated CR for ACSDs
    # ed_crudert = tcr_ed10, # their calculated CR for ed visits
    # pcp_crudert = tcr_pc10 # their calculted CR for pcp visits
  ) %>% 
  mutate_if(is.double, funs(if_else(. %in% c(-99, -999), NA_real_, .))) %>% 
  replace_na(list(pcpvt_fqhc = 0, pcpvt_rhc = 0)) %>% 
  mutate(
    pcpvt = partb_pcp + pcpvt_rhc + pcpvt_fqhc, # total pcp visits for part B and OTP
    physicians = pcphys + specialist + obgyn, # + famprac + internist,
    pa = pa_ob + pa_pcp + pa_spec + pa_oth,
    obstets = obgyn + pa_ob + cnm_fte,
    allpcp = pcphys + pa_pcp, # + famprac + internist,
    allpcpnp = pcphys + np_fte + pa_pcp, # + famprac + internist,
    specs = specialist + pa_spec + pa_oth,
    allimg = img_ob + img_pcp + img_spec
  )
```

### Crosswalk from 2010 tracts to 2010 zctas

```{r}
dart10_all3 <- dart10_all2 %>% 
  left_join(nyc_zcta_xwalk, by = "tract10") %>% 
  mutate_if(is.double, funs(. * afact)) %>% 
  select(-afact) %>% 
  group_by(zcta2010) %>% 
  summarise_if(is.double, sum, na.rm = T) %>% 
  mutate(year = 2010)

```


### Create 2010 zcta-level rates/ratios

```{r}
dart10_all4 <- dart10_all3 %>% 
  mutate(
    # add zcta-level doc ratios
    physician_rat = totpop / na_if(physicians, 0),
    pa_rat = totpop / na_if(pa, 0),
    np_rat = totpop / na_if(np_fte, 0),
    cnm_rat = fem15_44 / na_if(cnm_fte, 0),
    allpcp_rat = totpop / na_if(allpcp, 0),
    allpcpnp_rat = totpop / na_if(allpcpnp, 0),
    specs_rat = totpop / na_if(specs, 0),
    obstets_rat = fem15_44 / na_if(obstets, 0),
    obgyn_rat = fem15_44 / na_if(obgyn, 0),
    img_rat = totpop / na_if(allimg, 0),

    # add zcta-level visit rates
    acscd_rt = medicare_acscd / na_if(hospdenom, 0),
    pcpvt_rt = pcpvt / na_if(pcpdenom, 0),
    edvt_rt = edvt / na_if(pcpdenom, 0),

    phys_p1000 = (physicians/   na_if(totpop, 0))*1000,
    pcphys_p1000 = (pcphys / na_if(totpop, 0))*1000,
    pa_p1000 = (pa / na_if(totpop, 0))*1000,
    allpcp_p1000 = (allpcp /   na_if(totpop, 0))*1000,
    allpcpnp_p1000 = (allpcpnp /   na_if(totpop, 0))*1000,
    specs_p1000 = (specs / na_if(totpop, 0))*1000,
    obstets_p1000 = (obstets / na_if(totpop, 0))*1000,
    obgyn_p1000 = (obgyn / na_if(totpop, 0))*1000,
    img_p1000 = (allimg / na_if(totpop, 0))*1000,
    np_p1000 = (np_fte / na_if(totpop, 0))*1000
  )

```

# 1999 Data

### Import all files, filter to nyc, clean names, and merge together

```{r}
dart99_all <- c("../dropbox/capstone/data_inter/zcta99_zcta2010.csv", 
            "../dropbox/capstone/data_inter/cms99_zcta2010.csv") %>% 
  map(~read_csv(.x, col_types = cols(zcta2010 = "c")) %>% 
        janitor::clean_names()) %>% 
  reduce(left_join, by = "zcta2010")

# ## bring in 99 data
# raw99 <- read.csv("../dropbox/capstone/zcta99_zcta2010.csv", as.is = TRUE)
# names(raw99) <- names(raw99) %>% str_to_lower()
# 
# cms_pcsa <-read.csv("../dropbox/capstone/cms99_zcta2010.csv", as.is = T)
# names(cms_pcsa) <- names(cms_pcsa) %>% str_to_lower()
```


### Create all basic variables and rates/ratios 
### (already at 2010 zcta-level)

```{r}
dart99_all2 <- dart99_all %>%
  transmute(
    zcta2010 = zcta2010, 
    year = 1999,
    pcpvt = zvt_pc99, # part B & outpatient file
    ambvt = zvt_am99, # part B & outpatient file
    edvt = zvt_ed99, # total #ED visits per bene per day by Med Part B bene resident in ZCTA
    pcpdenom = zvden_99, 
    eddenom = zhden_99, # hospitalization denom- not sure for ED
    totbene = zbene_n,
    pcphys = zg_doc,
    specialist = zs_doc,
    obgyn = zo_doc,
    nonfedpcp = zg_nfeddoc,
    nonfedspec = zs_nfeddoc,
    nonfedob = zo_nfeddoc,
    img_pcp = zg_img,
    img_spec = zs_img,
    img_ob = zo_img,
    pa = zpa_tot, # total PAs
    pa_pcp = zpa_pc,
    pa_spec = zpa_spec,
    pa_ob = zpa_obgyn,
    pa_oth = zpa_oth,
    fqhcs = zfqhc,
    fem15_44 = zp012030 + zp012031 + zp012032 + zp012033 + zp012034 + zp012035 + zp012036 + zp012037 + zp012038,
    totpop = zage00_14 + zage15_64 + zage65_up
  ) %>% 
  mutate(
    physicians = pcphys + specialist + obgyn,
    obstets = obgyn + pa_ob,
    allpcp = pcphys + pa_pcp,
    specs = specialist + pa_spec + pa_oth,
    allimg = img_ob + img_pcp + img_spec,
    # PCSA-level doc ratios
    physician_rat = totpop / na_if(physicians, 0),
    pa_rat = totpop / na_if(pa, 0),
    allpcp_rat = totpop / na_if(allpcp, 0),
    specs_rat = totpop / na_if(specs, 0),
    obstets_rat = fem15_44 / na_if(obstets, 0),
    obgyn_rat = fem15_44 / na_if(obgyn, 0),
    img_rat = totpop / na_if(allimg, 0),
    # PCSA-level visit rates
    pcpvt_rt = (pcpvt / na_if(pcpdenom, 0))*100,
    edvt_rt = (edvt / na_if(eddenom, 0))*1000,

    # new99 does not include internist and famprac, separately, but those are included in total physicians
    # no np in pcp, no cnm in obstets for new99 
    phys_p1000 = (physicians / na_if(totpop, 0))*1000,
    pcphys_p1000 = (pcphys / na_if(totpop, 0))*1000,
    pa_p1000 = (pa / na_if(totpop, 0))*1000,
    allpcp_p1000 = (allpcp / na_if(totpop, 0))*1000,
    specs_p1000 = (specs / na_if(totpop, 0))*1000,
    obstets_p1000 = (obstets / na_if(totpop, 0))*1000,
    obgyn_p1000 = (obgyn / na_if(totpop, 0))*1000,
    img_p1000 = (allimg / na_if(totpop, 0))*1000
  )

```

# Combine 1999 & 2010 data, add gentrification status, export clean file

```{r}
gent_status <- read_csv("../dropbox/capstone/data_inter/zcta_gent_xwalk.csv", col_types = "cc") %>% 
  mutate(gent_status = ordered(gent_status, levels = c("Non-Gentrifying", "Gentrifying", "High Income")))

dart_all_years <- bind_rows(dart99_all2, dart10_all4) %>% 
  left_join(gent_status, by = "zcta2010") %>% 
  select(zcta2010, year, gent_status, everything())

write_csv(dart_all_years, "../dropbox/capstone/data_inter/dart_clean.csv")

```



