---
title: "Data & Methods"
output:
  github_document: default
  html_notebook: default
date: '`r Sys.Date()`'
---

```{r, message=FALSE}
library(tidyverse)
library(sf)
library(units)

options(scipen = 20)
```


## ZCTAs

```{r, results='hide'}
zcta2010 <- st_read("../dropbox/capstone/shapefiles/nyc_zcta2010/nyc_zcta2010.shp", "nyc_zcta2010", 
                    stringsAsFactors = FALSE)

zcta_vars <- feather::read_feather("../dropbox/capstone/data_clean/all_data.feather") %>% 
  transmute(zcta2010 = zcta2010,
            pop = pop_tot_2010, 
            inc90 = avg_inc_adj_1990,
            chg_rent = (avg_rent_adj_2010 - avg_rent_adj_1990) / avg_rent_adj_1990)
```


```{r}
dropped_zctas <- anti_join(zcta2010, zcta_vars, by = "zcta2010") %>% nrow()
```

There are `r dropped_zctas` ZCTAs excluded because their population was less than 2,000, leaving `r nrow(zcta_vars)` for our analysis.

```{r}
zcta_stats <- zcta2010 %>% 
  inner_join(zcta_vars, by = "zcta2010") %>% 
  mutate(area = st_area(.) %>% set_units(mi^2)) %>% 
  summarize_at(vars(area, pop), funs(mean, median, IQR, sd))
```

Of the ZCTAs we include, the median size is `r round(zcta_stats$area_median, 2)` and the median population is `r round(zcta_stats$pop_median, 0)`


## ZCTA Gentrification stats

```{r}
gent_stats <- zcta_vars %>% 
  summarise(inc90_40th = quantile(inc90, .4),
            rent_chg_med = median(chg_rent))
```

For our zcta-level gentrification definition:

40th percentile zcta for household income was `r round(gent_stats$inc90_40th, 0)` (2016$)
Median percent change in average gross rent between 1990 and 2010 was `r round(gent_stats$rent_chg_med * 100, 1)`%


## PCSAs

```{r}
# pcsa2010 <- st_read("../dropbox/capstone/shapefiles/pcsav3_1shapefiles/uspcsav31_HRSA.shp", "uspcsav31_HRSA", 
#                     stringsAsFactors = FALSE)
# 
# (pcsa_area <- pcsa2010 %>% 
#   mutate(area = st_area(.) %>% set_units(mi^2)) %>% 
#   summarize_at(vars(area), funs(mean, median, IQR, sd)))
```


## Tracts

```{r, results='hide'}
tract2010 <- st_read("../dropbox/capstone/shapefiles/nyct2010_17a/nyct2010.shp", "nyct2010", 
                    stringsAsFactors = FALSE)
tract_stats <- tract2010 %>% 
  mutate(area = st_area(.) %>% set_units(mi^2)) %>% 
  summarize_at(vars(area), funs(mean, median, IQR, sd))
```

The median 2010 census tract is `r round(tract_stats$median, 3)` square miles.
